Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-10 foreground_delay_spectrum_modeling.tex
From f0b7176 on branch danny2: =============================
(*) Made few minor edits and removed some commented sentences

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{ Delay spectrum power spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30. At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set. To get the the power spectrum each baseline is Fourier transformed into delay space ($y$-axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$-axis). The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

@@

\subsection{Foreground Model}\label{sec:foreground}
%todo
The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon. As we have seen above, foreground objects contaminate the region inside the ``wedge'' with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} --i.e. within the white lines of Figure \ref{fig:sim_data}-- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively. A value of zero indicates when the ratio $\rho=1$. \label{fig:data-sim-ratio}}
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Probability densities and cumulative distributions are shown in black and gray respectively. Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty. Also indicated in percentages are confidence levels estimated conservatively for such a consistency. They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. \label{fig:data-sim-ratio-snr}}
\end{figure}

@@

%The foreground signatures we henceforth identify in the delay spectrum are characterized as arising out of the aforementioned factors. Due to numerous combinations of such factors, we use unique annotations to characterize these signatures.

@@

\subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}
The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length. Therefore, the distribution of objects within the primary field of view depends mainly on how the power--pattern maps to delay space for each antenna pair. Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the pointing center. In the {\it zenith} pointing compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the main lobe of the power--pattern is centered at zenith.

@@

%\section{Discussion}\label{sec:discussion}

%\section{Conclusion}\label{sec:conclusion}

>>>>

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{ Delay spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30. At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set. To get the the power spectrum each baseline is Fourier transformed into delay space ($y$--axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$--axis). The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum objects are physically limited to the region below this line. The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the power--pattern and the sky at that instant. \label{fig:fhd_data}}
\end{figure}

@@

\subsection{Foreground Model}\label{sec:foreground}

The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non-zero response all the way to the horizon. As we have seen above, foreground objects contaminate the region inside the ``wedge'' with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviations of $\log_{10}\rho$ are 0.29 and 0.27 respectively. A value of zero indicates when the ratio $\rho=1$. \label{fig:data-sim-ratio}}
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_90.0K_185.0_MHz_30.7_MHz_bnw2.0.eps}
% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Probability densities and cumulative distributions are shown in black and gray respectively. Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty. Also indicated in percentages are confidence levels estimated conservatively for such a consistency. They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. \label{fig:data-sim-ratio-snr}}
\end{figure}

@@

@@

\subsubsection{Compact Emission in Power--Pattern Main Lobe}\label{sec:csm-main-lobe}

The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length. Therefore, the distribution of objects within the primary field of view depends mainly on how the power--pattern maps to delay space for each antenna pair. Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the pointing center. In the {\it zenith} pointing compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the main lobe of the power--pattern is centered at zenith.

@@
